---
title: "R Notebook"
output: pdf_notebook
---

set directory and read in dataset
```{r}
setwd('./data')
filename = 'MoothaData.txt';
mootha  = read.table(filename)
```

Install the packages if have not been done yet, ignore if have done so

install.packages("psych",dependencies=TRUE)

separate data by its feature, in our case, it's whether diabetes or not
```{r}
g1 = 1:17
g2 = 18:34
s1 = sample(g1, 2, replace=FALSE)
s2 = sample(g2, 2, replace=FALSE)
sm_sample = c(s1,s2)
mootha[,sm_sample]
```
```{r}
library(psych)
```

Function for plotting the matrix
We only include four person's data by randomly choosing from the two groups
```{r}
pairs.panels(mootha[,sm_sample], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```
```{r}
library(tidyr)
library(dplyr)
```
Here we want to find the summary of the dataset's stats
First use the gather function from tidyr

Then save the stats as a table, including mean, standard deviation, max and min of each person's gene expression
```{r}
mootha_long <- gather(mootha, factor_key=TRUE)

summarydf <- mootha_long %>% group_by(key) %>%
  summarize(mean= mean(value), sd= sd(value), max = max(value),min = min(value))

summarydf
```

```{r}
write.csv(summarydf, filename='stats_summary.csv')
```

Redo and find the mean of each stats by groups, save them as a dataframe
```{r}
g1_stat = apply(summarydf[g1,2:5],2,mean)
g2_stat = apply(summarydf[g2,2:5],2,mean)

g1_stat
g2_stat

group_stat = data.frame(g1_stat,g2_stat)
```
```{r}
g1_stat['mean']
```



